High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-04-23 notebooks/mnist.qmd
working my boi :sunglasses:

```{julia}
# Model parameters:
epochs = 100
batch_size = minimum([Int(round(n_obs/10)), 128])
n_hidden = 128
activation = Flux.swish
builder = MLJFlux.@builder Flux.Chain(

@@

    Dense(n_hidden, n_out),
)
# builder = MLJFlux.Short(n_hidden=n_hidden, dropout=0.1, σ=activation)
# builder = MLJFlux.MLP(
#     hidden=(
#         n_hidden,
#         n_hidden,
#         n_hidden,
#     ), 
#     σ=activation
# )
α = [1.0,1.0,1e-2]

# Simple MLP:
mlp = NeuralNetworkClassifier(
    builder=builder, 
    epochs=epochs,
    batch_size=batch_size,
)

@@

# Deep Ensemble:
mlp_ens = EnsembleModel(model=mlp, n=50)
```

```{julia}
cov = .95
conf_model = conformal_model(jem; method=:simple_inductive, coverage=cov)
mach = machine(conf_model, X, labels)
fit!(mach)
M = ECCCo.ConformalModel(mach.model, mach.fitresult)
```

```{julia}
if mach.model.model isa JointEnergyModels.JointEnergyClassifier
    jem = mach.model.model.jem
    n_iter = 200
    _w = 1500
    plts = []
    neach = 10
    for i in 1:10
        x = jem.sampler(jem.chain, jem.sampling_rule; niter=n_iter, n_samples=neach, y=i)
        plts_i = []
        for j in 1:size(x, 2)
            xj = x[:,j]
            xj = reshape(xj, (n_digits, n_digits))
            plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
        end
        plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
        plts = [plts..., plt]
    end
    plt = Plots.plot(plts..., size=(_w,_w), layout=(10,1))
    display(plt)
end
```

@@

# Set up search:
factual_label = 9
x = reshape(counterfactual_data.X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 4
factual = predict_label(M, counterfactual_data, x)[1]
γ = 0.5
T = 250

@@

# Generate counterfactual using generic generator:
generator = GenericGenerator()
ce_wachter = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)

generator = GreedyGenerator(η=η)
ce_jsma = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
)

@@

# Generate counterfactual using ECCCo generator:
generator = ECCCoGenerator(
    λ=λ, 
    temp=temp, 
)
ce_conformal = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)

>>>>

```{julia}
# Model parameters:
epochs = 100
batch_size = minimum([Int(round(n_obs/10)), 128])
n_hidden = 128
activation = Flux.relu
builder = MLJFlux.@builder Flux.Chain(

@@

    Dense(n_hidden, n_out),
)
# builder = MLJFlux.Short(n_hidden=n_hidden, dropout=0.1, σ=activation)
# builder = MLJFlux.MLP(
#     hidden=(
#         n_hidden,
#         n_hidden,
#         n_hidden,
#     ), 
#     σ=activation
# )
α = [1.0,1.0,5e-3]

# Simple MLP:
mlp = NeuralNetworkClassifier(
    builder=builder, 
    epochs=epochs,
    batch_size=batch_size,
    finaliser=x -> x,
    loss=Flux.Losses.logitcrossentropy,
)

# Deep Ensemble:
mlp_ens = EnsembleModel(model=mlp, n=5)

@@

# Deep Ensemble of Joint Energy Models:
jem_ens = EnsembleModel(model=jem, n=5)
```

```{julia}
cov = .95
conf_model = conformal_model(jem_ens; method=:adaptive_inductive, coverage=cov)
mach = machine(conf_model, X, labels)
fit!(mach)
M = ECCCo.ConformalModel(mach.model, mach.fitresult)
```

```{julia}
if mach.model.model isa JointEnergyClassifier
    sampler = mach.model.model.jem.sampler
else
    K = length(counterfactual_data.y_levels)
    input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
    𝒟x = Uniform(extrema(counterfactual_data.X)...)
    𝒟y = Categorical(ones(K) ./ K)
    sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
end
opt = ImproperSGLD()
f(x) = logits(M, x)

n_iter = 200
_w = 1500
plts = []
neach = 10
for i in 1:10
    x = sampler(f, opt; niter=n_iter, n_samples=neach, y=i)
    plts_i = []
    for j in 1:size(x, 2)
        xj = x[:,j]
        xj = reshape(xj, (n_digits, n_digits))
        plts_i = [plts_i..., Plots.heatmap(rotl90(xj), axis=nothing, cb=false)]
    end
    plt = Plots.plot(plts_i..., size=(_w,0.10*_w), layout=(1,10))
    plts = [plts..., plt]
end
plt = Plots.plot(plts..., size=(_w,_w), layout=(10,1))
display(plt)

```

@@

# Set up search:
factual_label = 4
x = reshape(counterfactual_data.X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target = 5
factual = predict_label(M, counterfactual_data, x)[1]
γ = 0.9
T = 100

@@

# Generate counterfactual using generic generator:
generator = GenericGenerator()
ce_wachter = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)

generator = GreedyGenerator(η=η)
ce_jsma = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)

@@

# Generate counterfactual using ECCCo generator:
generator = ECCCoGenerator(
    λ=λ, 
    temp=temp, 
    opt=Flux.Optimise.Adam(0.01),
)
ce_conformal = generate_counterfactual(
    x, target, counterfactual_data, M, generator; 
    decision_threshold=γ, max_iter=T,
    initialization=:identity,
    converge_when=:generator_conditions,
)
